---
title: Scrabble two letter words
output:
 html_document:
      theme: flatly
      highlight: textmate
---

```{r markdown, echo = FALSE, warning = FALSE}
library(knitr)
library(devtools)
opts_chunk$set(echo = FALSE, eval = TRUE, cache = FALSE, comment = "")
```

```{r setup, warning = FALSE}
library(glue)
library(magrittr)
```

```{r}
words <- c("ae", "ah", "ai", "al", "ar", "aw", "ay", "ba", "bo", "da", "de", "di", "ea", "ef", "eh", "el", "es", "et", "fa", "fe", "gi", "gu", "ha", "ho", "id", "io", "ka", "ki", "ko", "la", "li", "lo", "ob", "od", "oe", "oh", "oi", "os", "ou", "ow", "oy", "pa", "pe", "po", "si", "te", "ti", "ug", "uh", "ur", "ut", "wo", "ya", "ye", "yo", "yu")
```

There are 124 two letter words

```{r}
words_common <- c("am", "an", "as", "at", "be", "by", "do", "go", "he", "hi", "if", "in", "is", "it", "me", "my","no", "of", "on", "or", "ox", "pi", "so", "to", "up", "us", "we")
```

**`r length(words_common)` are common words**
```{r}
words_common_string <- collapse(words_common, sep = " ") %>% cat()
```

```{r}
words_abbreviations <- c("ab", "ad", "ag", "ax", "bi", "ed", "ma", "mo", "ne", "op", "ta")
```

**`r length(words_abbreviations)` are abbreviations**
```{r}
words_abbreviations_string <- collapse(words_abbreviations, sep = " ") %>% cat()
```

**ab** = *abdominal muscle*
**ad** = *advertisement*
**ag** = *agriculture*
**ax** = *axe*
**bi** = *bisexual*
**ed** = *editor*
**ma** = *mother*
**mo** = *moment*  
**ne** = *formally known as*
**op** = *operation*
**ta** = *thanks*

```{r}
words_prefix <- c("er", "ex", "re", "un" )
```

**`r length(words_prefix)` are prefixes or suffixes**
```{r}
words_prefix_string <- collapse(words_prefix, sep = " ") %>% cat()
```

```{r}
words_restricted <- c("ch", "ja", "jo", "qi", "za", "zo")
```

**5 letters, c, j, q, v, z only contain 1 or 2 two letter words**

* **v** has no two letter words
* **c, q** has one two letter word
* **j, z** has two two letter words ending in either a or o

```{r}
words_restricted_string <- collapse(words_restricted, sep = " ") %>% cat()
```

**ch** = *old word meaning to lengthen*
**ja** = *South African word for yes*
**jo** = *Scots word for sweetheart*
**qi** = *Chinese life force*
**za** = *pizza*  
**zo** = *Tibetan yak x cow hybrid*

```{r}
words_vowels <- c("em", "en", "mi", "mu", "na", "nu", "om", "um")
```

**2 letters m and n can have most vowels before or after them** (*except im and ni*)

```{r}
words_vowels_string <- collapse(words_vowels, sep = " ") %>% cat()
```

**em** = *unit of measurement*
**en** = *unit of measurement*
**mi** = *musical note*
**mu** = *Greek letter*
**na** = *Scots word for no*
**nu** = *Greek letter*  
**om** = *Hindu syllable*
**um** = *sound of hesitation*

```{r}
words_x_vowels <- c("xi", "xu")
```

**1 letter x has 3 vowels before and 2 vowels after it**

```{r}
words_x_vowels_string <- collapse(words_x_vowels, sep = " ") %>% cat()
```

**xi** = *Greek letter*
**xu** = *Vietnamese currency*

```{r}
words_duplicate <- c("aa", "ee", "oo", "mm")
```

**`r length(words_duplicate)` contain duplicate letters**
```{r}
words_duplicate_string <- collapse(words_duplicate, sep = " ") %>% cat()
```

**aa** = *rough volcanic rock*
**ee** = *scots word for eye*
**oo** = *scots word for wool*
**mm** = *sound of enjoyment*

```{r}
words_consonants <- c("fy", "hm", "ky", "ny", "sh", "st")
```

**`r length(words_consonants)` only contain consonants**
```{r}
words_consonants_string <- collapse(words_consonants, sep = " ") %>% cat()
```

**fy** = *sound of disapproval*
**hm** = *sound of doubt*
**ky** = *scots word for cows*
**ny** = *near*
**sh** = *sound requesting quiet*  
**st** = *sound to attract attention*

There are `r length(words)` words remaining
```{r}
words_string <- collapse(words, sep = " ") %>% cat()
```
