---
title: Scrabble two letter words
output:
 html_document:
      theme: flatly
      highlight: textmate
---

```{r markdown, echo = FALSE, warning = FALSE}
library(knitr)
library(devtools)
opts_chunk$set(echo = FALSE, eval = TRUE, cache = FALSE, comment = "")
```

```{r setup, message = FALSE}
library(purrr)
library(glue)
library(here)
library(magrittr)
```

```{r}
rules <- list()
```

```{r}
words <- list(
  name = "Remaining words",
  before  = NA_character_,
  consonant = NA_character_,
  after = NA_character_,
  words = c("aa", "ab", "ad", "ae", "ag", "ah", "ai", "al", "am", "an", "ar", "as", "at", "aw", "ax", "ay", "ba", "be", "bi", "bo", "by", "ch", "da", "de", "di", "do", "ea", "ed", "ee", "ef", "eh", "el", "em", "en", "er", "es", "et", "ex", "fa", "fe", "fy", "gi", "go", "gu", "ha", "he", "hi", "hm", "ho", "id", "if", "in", "io", "is", "it", "ja", "jo", "ka", "ki", "ko", "ky", "la", "li", "lo", "ma", "me", "mi", "mm", "mo", "mu", "my", "na", "ne", "no", "nu", "ny", "ob", "od", "oe", "of", "oh", "oi", "om", "on", "oo", "op", "or", "os", "ou", "ow", "ox", "oy", "pa", "pe", "pi", "po", "qi", "re", "sh", "si", "so", "st", "ta", "te", "ti", "to", "ug", "uh", "um", "un", "up", "ur", "us", "ut", "we", "wo", "xi", "xu", "ya", "ye", "yo", "yu", "za", "zo")
)
```

## Common words
```{r}
rule <- list(
  name = "Common words",
  before  = NA_character_,
  consonant = NA_character_,
  after = NA_character_,
  words = c("am", "an", "as", "at", "be", "by", "do", "go", "he", "hi", "if", "in", "is", "it", "me", "my", "no", "of", "on", "or", "ox", "so", "to", "up", "us", "we")
)
```

```{r, rule_formatting}
# get word list
word_list <- rule %>% pluck("words")
# add rule to rules list
rules <- rules %>% append(list(rule))
# remove word list from words
words[["words"]] <- words[["words"]] %>% 
  discard(words[["words"]] %in% word_list)
# print word list
collapse(word_list, sep = " ") %>% cat()
```

## All vowels

### **aeiou** vowels before consonants **ns**
```{r}
rule <- list(
  name = "aeiou vowels before consonants ns",
  before  = "aeiou",
  consonant = "ns",
  after = NA_character_,
  words = c("an", "en", "in", "on", "un", "as", "es", "is", "os", "us")
)
```

```{r ref.label='rule_formatting'}
```

**en** = *unit of measurement*
**es** = *letter s*
**os** = *mouthlike opening*
**un** = *one*

### **aeiou** vowels after consonant **m**
```{r}
rule <- list(
  name = "aeiou vowels after consonant m",
  before  = NA_character_,
  consonant = "m",
  after = "aeiou",
  words = c("ma", "me", "mi", "mo", "mu")
)
```

```{r ref.label='rule_formatting'}
```

**ma** = *abbreviation for mother*
**mi** = *musical note*
**mo** = *abbreviation for moment*
**mu** = *Greek letter*

## **aeo** vowels

### **aeo** vowels surround themselves and their adjacent consonants
```{r}
rule <- list(
  name = "aeo vowels surround themselves and their adjacent consonants",
  before  = "aeo",
  consonant = NA_character_,
  after = "aeo",
  words = c("aa", "ab", "ba", "de", "ed", "ee", "ef", "fe", "on", "no", "oo", "op", "po")
)
```

```{r ref.label='rule_formatting'}
```

**aa** = *rough volcanic rock*
**ab** = *abbreviaion for abdominal muscle*
**ba** = *Ancient Eygptian symbol for soul*
**de** = *prefix of or from*  
**ed** = *abbreviation for editor*
**ee** = *scots word for eye*
**ef** = *letter f*
**fe** = *fee*
**po** = *bedpan*
**oo** = *scots word for wool*
**op** = *abbreviation for operation*

### **aeo** vowels surround consonants **dhmn**
```{r}
rule <- list(
  name = "aeo vowels surround consonants dhmn",
  before  = "aeo",
  consonant = "dhmn",
  after = "aeo",
  words = c("ad", "ed", "od", "da", "de", "do", "ah", "eh", "oh", "ha", "he", "ho", "am", "em", "om", "ma", "me", "mo", "an", "en", "on", "na", "ne", "no")
)
```

```{r ref.label='rule_formatting'}
```

**ad** = *abbreviation for advertisement*
**ah** = *sound of surprise*
**da** = *Burmese knife*
**de** = *prefix of or from*
**ed** = *abbreviation for editor*  
**eh** = *sound of surprise*
**em** = *unit of measurement*
**ha** = *sound of laughter*
**ho** = *sound of laughter*
**ma** = *abbreviation for mother*  
**mo** = *abbreviation for moment*
**na** = *Scots word for no*
**ne** = *abbrevation for formally known as*
**od** = *mystical force*
**oh** = *sound of surprise*  
**om** = *Hindu syllable*

### **aeo** vowels before consonants **rx**
```{r}
rule <- list(
  name = "aeo vowels before consonants rx",
  before  = "aeo",
  consonant = "rsx",
  after = NA_character_,
  words = c("ar", "er", "or", "ax", "ex", "ox")
)
```

```{r ref.label='rule_formatting'}
```

**ar** = *letter r*
**ax** = *axe*
**er** = *sound of hesitation*
**ex** = *excluding*

### **aeo** vowels after consonants **bpty**
```{r}
rule <- list(
  name = "aeo vowels after consonants bpty",
  before  = NA_character_,
  consonant = "bpty",
  after = "aeo",
  words = c("ba", "be", "bo", "pa", "pe", "po", "ta", "te", "to", "ya", "ye", "yo")
)
```

```{r ref.label='rule_formatting'}
```

**ba** = *Ancient Eygptian symbol for soul*
**bo** = *sound to startle*
**pa** = *abbreviation for father*
**pe** = *Hebrew letter*
**po** = *bedpan*  
**ta** = *abbreviation for thanks*
**te** = *musical note*
**ya** = *you*
**ye** = *the*
**yo** = *sound to attract attention* 

## **iu** vowels

### **iu** vowels before consonant **t**
```{r}
rule <- list(
  name = "iu vowels before consonants t",
  before  = "iu",
  consonant = "t",
  after = NA_character_,
  words = c("it", "ut")
)
```

```{r ref.label='rule_formatting'}
```

**ut** = *musical note*

### **iu** vowels after consonants **gx**
```{r}
rule <- list(
  name = "iu vowels after consonants gx",
  before  = NA_character_,
  consonant = "gx",
  after = "iu",
  words = c("gi", "gu", "xi", "xu")
)
```

```{r ref.label='rule_formatting'}
```

**gi** = *martial arts white clothing*
**gu** = *shetland violin*
**xi** = *Greek letter*
**xu** = *Vietnamese currency*

## **aio** vowels

### **aio** vowels surround consonant **d**
```{r}
rule <- list(
  name = "aio vowels surround consonant d",
  before  = "aio",
  consonant = "d",
  after = "aio",
  words = c("ad", "da", "di", "id", "do", "od")
)
```

```{r ref.label='rule_formatting'}
```

**ad** = *abbreviation for advertisement*
**da** = *Burmese knife*
**di** = *gods*
**id** = *unconscious instinct*
**od** = *mystical force*

### **aio** vowels after consonants **bhklpt**
```{r}
rule <- list(
  name = "aio vowels after consonants bhklpt",
  before  = NA_character_,
  consonant = "bhklmpt",
  after = "aio",
  words = c("ba", "bi", "bo", "ha", "hi", "ho", "ka", "ki", "ko", "la", "li", "lo", "pa", "pi", "po", "ta", "ti", "to")
)
```

```{r ref.label='rule_formatting'}
```

**ba** = *ancient Eygptian symbol for soul*
**bi** = *abbreviation for bisexual*
**bo** = *sound to startle*
**ha** = *sound of laughter*
**ho** = *sound of laughter*  
**ka** = *Egyptian spirit*
**ki** = *Japanese martial art*
**ko** = *New Zealand digging tool*
**la** = *musical note*
**li** = *Chinese measurement of distance*
**lo** = *look*  
**pa** = *abbreviation for father*
**pi** = *Hebrew letter*
**po** = *bedpan*  
**ta** = *abbreviation for thank you*
**ti** = *musical note*

## Difficult consonants

### **c** is followed by consonant **h**
```{r}
rule <- list(
  name = "c is only followed by consonant h",
  before  = NA_character_,
  consonant = "c",
  after = "h",
  words = c("ch")
)
```

```{r ref.label='rule_formatting'}
```

**ch** = *old word meaning to lengthen*

### **jz** are followed by vowels **ao**
```{r}
rule <- list(
  name = "jz are only followed by vowels ao",
  before  = NA_character_,
  consonant = "jz",
  after = "ao",
  words = c("ja", "jo", "za", "zo")
)
```

```{r ref.label='rule_formatting'}
```

**ja** = *South African word for yes*
**jo** = *Scots word for sweetheart*
**za** = *pizza*
**zo** = *Tibetan yak x cow hybrid*

### **q** is followed by vowel **i**
```{r}
rule <- list(
  name = "q is only followed by vowel i",
  before  = NA_character_,
  consonant = "q",
  after = "i",
  words = c("qi")
)
```

```{r ref.label='rule_formatting'}
```

**qi** = *Chinese life force*

### **r** is followed by vowel **e**
```{r}
rule <- list(
  name = "r is followed only by vowel e",
  before  = NA_character_,
  consonant = "r",
  after = "e",
  words = c("re")
)
```

```{r ref.label='rule_formatting'}
```

**re** = *prefix meaning concerning*

### **v** is not surrounded by any vowel or consonant
```{r}
rule <- list(
  name = "v is not surrounded by any vowel or consonant",
  before = NA_character_,
  consonant = "v",
  after = NA_character_,
  words = NA_character_
)
```

### **x** is preceded by vowels **aeo** and followed by vowels **iu**
```{r}
rule <- list(
  name = "x is preceded by vowels aeo and followed by vowels iu",
  before = "aeo",
  consonant = "x",
  after = "iu",
  words = c("ax", "ex", "ox", "xi", "xu")
)
```

```{r ref.label='rule_formatting'}
```

**ax** = *axe*
**ex** = *excluding*
**xi** = *Greek letter*
**xu** = *Vietnamese currency*

## Vowels only
```{r}
rule <- list(
  name = "Vowels only",
  before  = NA_character_,
  consonant = NA_character_,
  after = NA_character_,
  words = c("aa", "ae", "ai", "ea", "ee", "io", "oe", "oi", "oo", "ou")
)
```

```{r ref.label='rule_formatting'}
```

**aa** = *rough volcanic rock*
**ae** = *scots word for one*
**ai** = *three-toed sloth*
**ea** = *river*
**ee** = *scots word for eye*
**io** = *moth*
**oe** = *grandchild*  
**oi** = *sound to attract attention*
**oo** = *scots word for wool*
**ou** = *man*

## Consonants only
```{r}
rule <- list(
  name = "Consonants only",
  before  = NA_character_,
  consonant = NA_character_,
  after = NA_character_,
  words = c("by", "ch", "fy", "hm", "ky", "mm", "my", "ny", "sh", "st")
)
```

```{r ref.label='rule_formatting'}
```

**ch** = *old word meaning to lengthen*
**fy** = *sound of disapproval*
**hm** = *sound of doubt*
**ky** = *Scots word for cows*
**mm** = *sound of enjoyment*
**ny** = *near*  
**sh** = *sound requesting quiet*
**st** = *sound to attract attention*

## Remaining words

There are `r length(words[["words"]])` words remaining

```{r}
rules <- rules %>% append(list(words))
collapse(words[["words"]], sep = " ") %>% cat()
```

**ag** = *abbreviation for agriculture*
**al** = *Asian shrub or tree*
**aw** = *all*
**ay** = *yes*
**el** = *letter l*
**et** = *ate*
**fa** = *musical note*
**nu** = *Greek letter*  
**ob** = *expression of opposition*
**ow** = *sound of pain*
**oy** = *grandchild*
**si** = *musical note*
**ug** = *hate*
**uh** = *sound of hesitation*
**um** = *sound of hesitation*  
**ur** = *sound of hesitation*
**wo** = *woe*
**yu** = *jade*

```{r echo = FALSE}
saveRDS(rules, here("two_letter_word_rules.rds"))
```

