---
output: html_document
---

```{r markdown, echo = FALSE, warning = FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, eval = TRUE, cache = FALSE, comment = "")
```

```{r setup, message = FALSE, warning = FALSE}
library(purrr)
library(dplyr)
library(tibble)
library(tidyr)
library(magrittr)
```

# Example lists

Lists can be viewed using str, limited by level or length, or interactively using listviewer::jsonedit

## Simple list

```{r}
simple_list <- list(a = "4a",
                    b = "4b",
                    c = "4c")
str(simple_list, max.level = 1, list.len = 3)
#listviewer::jsonedit(simple_list, mode = "view", width = 300, height = 200)
```

## List of lists

```{r}
list_of_lists <- list(one = list(a = "1a",
                                 b = "1b",
                                 c = "1c"),
                      two = list(a = "2a",
                                 b = "2b",
                                 c = "2c"),
                      three = list(a = "3a",
                                   b = "3b",
                                   c = "3c"),
                      four = list(a = "4a",
                                  b = "4b",
                                  c = "4c"))
str(list_of_lists, max.level = 2, list.len = 4)
#listviewer::jsonedit(list_of_lists, mode = "view", width = 300, height = 500)
```

## Nested list

```{r}
nested_list <- list(a = "1a",
                    b = "1b",
                    c = "1c",
                    two = list(a = "2a",
                               b = "2b",
                               c = "2c",
                               three = list(a = "3a",
                                            b = "3b",
                                            c = "3c",
                                            four = list(a = "4a",
                                                        b = "4b",
                                                        c = "4c"))))
str(nested_list, max.level = 4, list.len = 4)
#listviewer::jsonedit(nested_list, mode = "view", width = 300, height = 500)
```

# Subsetting

All these will return the value "4c", where a double bracket returns the component and single bracket a list

```{r}
simple_list[[3]]

list_of_lists[[4]][[3]]

nested_list[[4]][[4]][[4]][[3]]
nested_list[["two"]][["three"]][["four"]][["c"]]
nested_list$two$three$four$c
```

# purrr

map works best with the list of lists structure where it can extract one or more elements from each list

Extract function is from magrittr and equivalent to using `[`

```{r}
map(list_of_lists, "c")
# map(list_of_lists, extract, c("a", "b", "c"))
```

In the nested list structure map functions iterates through the first list element only in the list. In the nested list example, first drill down to the 3rd level using subsetting and then map will return NA for the first three elements as they are not lists, and "4c" from the fourth list element.

```{r}
map(nested_list[["two"]][["three"]], extract, "c")
```

Alternatively nested map functions can be used to iterate to the required list 

```{r}
nested_list %>% map(. %>% map(. %>% map(extract, "c")))
```

# Lists to tibbles and back

For a **simple list**, this can be done by subsetting to produce a wide format or using unlist and as_tibble to produce a long format

```{r}
tibble_wide <- simple_list %>% {
  tibble(
   a = simple_list[["a"]],
   b = simple_list[["b"]],
   c = simple_list[["c"]]
  )
}
tibble_long <- as_tibble(unlist(simple_list))
```

For a **list of lists** the map functions can be used. A { bracket is required to prevent list_of_lists being passed as the first arguement into tribble

```{r}
tibble_wide <- list_of_lists %>% {
  tibble(
   a = map_chr(., "a"),
   b = map_chr(., "b"),
   c = map_chr(., "c")
  )
}
```

tibble::enframe can be used to **convert a named list to a dataframe**, with two rows, one with the list element names and the second column with the lists as a list column.

If the list is not a named list the names can be first added from the list elements with purrr::set_names

```{r}
# col_names_vector <- map_chr(list_of_lists, "a")
tibble_list_column <- list_of_lists %>%
  # set_names(col_names_vector) %>% 
  enframe("level", "list_column")
```

The **list column can be converted back to a list_of_lists** using tibble::deframe

```{r}
list_of_lists <- deframe(tibble_list_column)
str(list_of_lists, max.level = 2, list.len = 4)
```

Individual list **elements within the list column can be extracted into their own columns** using the map functions, making a wide tibble

```{r}
tibble_wide <- tibble_list_column %>% 
  mutate(a = map_chr(list_column, "a"),
         b = map_chr(list_column, "b"),
         c = map_chr(list_column, "c"))
```

tidyr::unnest function can **convert the list column tibble into a long tibble format**

```{r}
tibble_long <- unnest(tibble_list_column)
```

The **tibble long format can be converted back to a tibble list column** using tidyr::nest function, grouped by the column that will produce the first level list names

```{r}
tibble_list_column <- tibble_long %>% 
  group_by(level) %>% 
  nest() %>% 
  ungroup()
```

[HERE](https://jennybc.github.io/purrr-tutorial/ls03_map-function-syntax.html)
