---
title: "Titanic survival machine learning analysis"
author: Graham French
date: '`r Sys.Date()`'
output:
 html_document:
      theme: flatly
      highlight: textmate
      toc: true
      toc_float:
        collapsed: true
        smooth_scroll: true
---

```{r markdown, echo = FALSE, warning = FALSE}
library(knitr)
library(devtools)
opts_chunk$set(echo = TRUE, eval = TRUE, cache = FALSE, warning = FALSE, comment = "")
```

```{r message = FALSE, warning = FALSE}
# Tidyverse
library(dplyr)
library(purrr)
library(tidyr)
library(readr)
library(stringr)
library(forcats)
library(janitor)
library(here)
library(magrittr)

# data
library(titanic)

# models
library(rpart)
library(rattle)
library(randomForest)
library(party)
```

## Titanic dataset

The titanic package contains both the training and test data. Columns names converted to lower case

```{r}
train <- titanic_train %>% 
  clean_names()
test <- titanic_test %>% 
  clean_names()
```

## Feature engineering

Training and test datasets combined to clean data and extract additional useful variables

```{r}
titanic <- bind_rows(train, test) 
```

### survived

Feature to predict in test dataset. Converted to factor

```{r }
titanic <- titanic %>% 
  mutate(survived = as_factor(as.character(survived)))

titanic %>% 
  tabyl(survived) %>% 
  adorn_pct_formatting(digits = 1)
```

### pclass

Three passenger class, feature kept as is, just converted to a factor

```{r}
titanic <- titanic %>% 
  mutate(pclass = as_factor(as.character(pclass)))

titanic %>% 
  tabyl(pclass) %>% 
  adorn_pct_formatting(digits = 1)
```

### name

Useful features extracted from the names were the titles and surnames used to determine family groups

**Title**

An additional variable extracted was the title from the names. This was taken from Vincent Broute's post [Titanic EDA & predictions attempt](https://www.kaggle.com/neveldo/titanic-eda-predictions-attempt/notebook)

```{r}
titanic <- titanic %>%
  mutate(title = str_extract(name, regex("([a-z]+\\.)", ignore_case = TRUE)),
         title = str_replace(title, "\\.", ""),
         title = as_factor(title))

# Group titles into refined groups
titanic <- titanic %>%
  mutate(title_bins = case_when
         (
           title %in% c("Capt", "Don", "Jonkheer", "Rev", "Mr") ~ "title_1",
           title %in% c("Col", "Dr", "Major", "Master") ~ "title_2",
           TRUE ~ "title_3"
          )
      ) %>% 
  mutate(title_bins = as_factor(title_bins))
         
titanic %>% 
  tabyl(title, title_bins) 
```

**surname**

```{r}
titanic <- titanic %>%
  mutate(surname = str_extract(name, regex("([a-z]+,)", ignore_case = TRUE)),
         surname = str_replace(surname, ",", ""))
```

### sex

Useful feature, keep as is, just convert to a factor

```{r}
titanic <- titanic %>% 
  mutate(sex = as_factor(sex))

titanic %>%
  tabyl(sex) %>% 
  adorn_pct_formatting(digits = 1)
```

### age

Age is not know for 263 passengers. Missing ages predicted using a decision tree

```{r}
# Training dataset
train_age <- titanic %>% 
  filter(!is.na(age))

# Run decision tree on training dataset with known ages
fit_age <- rpart(age ~ pclass + sex + sib_sp + parch + fare + embarked,
                 data = train_age,
                 method = "anova")

# Update missing ages with predicted ages
test_age <- titanic %>% 
  filter(is.na(age)) %>% 
  mutate(age = predict(fit_age, .))

# Recombine training and test datasets
titanic <- bind_rows(train_age, test_age)
```

[Vincent Broute analysis]((https://www.kaggle.com/neveldo/titanic-eda-predictions-attempt/notebook)) found that passengers 12 years old or less had a greater chance of survival and passengers 18-25 years old has a less chance so group age into these 3 groups

```{r}
titanic <- titanic %>% 
  mutate(age_bins = cut(age, breaks = c(0, 12, 18, 25, 80), 
                         include.lowest = TRUE,
                         labels = c("age_0-12", "age_12-18", "age_18-25", "age_25-80"),
                         ordered_factor = TRUE)) %>% 
  mutate(age_bins = fct_other(age_bins, keep = c("age_0-12", "age_18-25")))

titanic %>% 
  tabyl(age_bins) %>% 
  adorn_pct_formatting(digits = 1)
```

### sib_sp and parch

sib_sp (number of sibling or spouses aboard) and parch (number of parents or children aboard) can by combined to create a new **family_size** feature where family size is the number of family members a passenger is travelling with, adding up the siblings, spouses, parents and children, and adding one for the passenger themselves

```{r}
titanic <- titanic %>%
  mutate(family_size = sib_sp + parch + 1)
```

Analysis done by [Megan Risdal](https://www.kaggle.com/mrisdal/exploring-survival-on-the-titanic) shows that singletons or larger families above 4 members did not survive as much as passengers within a family of less than 4 members so group family size into these 3 groups

```{r}
titanic <- titanic %>%
    mutate(family_size_bins = cut(family_size, breaks = c(0, 1, 4, 11), 
                         include.lowest = TRUE,
                         labels = c("singleton", "family_size_1-4", "family_size_1-11"),
                         ordered_factor = TRUE))

titanic %>% 
  tabyl(family_size_bins) %>% 
  adorn_pct_formatting(digits = 1)
```

[Trevor Stephens](http://trevorstephens.com/kaggle-titanic-tutorial/r-part-4-feature-engineering/) suggested larger famlies would have had difficulty keeping together. To dentify each family, family size was combined with surname to create new **family_id** feature. This assumed that families with the same surname did not have the same family size

```{r}
titanic <- titanic %>%
  unite(family_id, surname, family_size, remove = FALSE)
```

Trevor combined passengers within small family units of 1-2 individuals

```{r}
titanic <- titanic %>%
  mutate(family_id = if_else(family_size <= 2, "small_family_unit", family_id))
```

### ticket

Feature not currently used in model

### fare

One third class passenger fare is unknown, replaced with the median price of third class fare ticket

```{r}
titanic %>% 
  group_by(pclass) %>% 
  summarise(avg_fare = median(fare, na.rm = TRUE))

titanic <- titanic %>%
  mutate(fare = replace_na(fare, 8.05))
```

### cabin

High number of missing values, feature not used in model

### embarked

Two passengers have an unknown embarked port. Analysis done by [Megan Risdal](https://www.kaggle.com/mrisdal/exploring-survival-on-the-titanic) shows that the median fare for the class both passengers coincides with them embarking at Charbourg. Feature converted to factor

```{r}
titanic <- titanic %>%
  mutate(embarked = if_else(embarked == "", "C", embarked))

titanic <- titanic %>% 
  mutate(embarked = as_factor(embarked))

titanic %>% 
  tabyl(embarked) %>% 
  adorn_pct_formatting(digits = 1)
```

**Export titanic dataset**

```{r}
titanic <- titanic %>% 
  select(passenger_id, 
         name, surname, title, title_bins,
         sex,
         age, age_bins,
         family_id, family_size, family_size_bins, sib_sp, parch,
         ticket, fare, cabin, pclass, embarked,
         survived) %>% 
  arrange(surname, family_id, -age) %T>% 
  write_csv(here("titanic_tidied_data.csv"))
```

## Training dataset

```{r}
# Training dataset
train_survived <- titanic %>% 
  filter(!is.na(survived))
```

## Logistic regression

Logistic regression to find the statsitically significant predictor features used by [Vincent Boute](https://www.kaggle.com/neveldo/titanic-eda-predictions-attempt/notebook)

```{r}
fit_survived <- glm(survived ~ title_bins + sex + age + family_size_bins + sib_sp + parch + fare + pclass + embarked,
                    data = train_survived,
                    family = binomial)
summary(fit_survived)
```

The prediction features statistically significant according to p-values were title_bins, age and pclass. Vincent also found the sib_sp was a significantly significant predictor feature

```{r}
fit_survived <- glm(survived ~ title_bins + age + pclass + sib_sp,
                    data = train_survived,
                    family = binomial)
summary(fit_survived)
```

## Models

### Decision Trees

[Trevor Stephen's series of titanic tutorials Part 3](http://trevorstephens.com/kaggle-titanic-tutorial/r-part-3-decision-trees/) tutorial calculated survival using decision trees, using the default control values and titanic features

```{r eval=FALSE, fig.width=8}
# Run decision tree on training dataset with known survival
fit_survived <- rpart::rpart(survived ~ title_bins + age + pclass + sib_sp,
                      data = train_survived,
                      method = "class",
                      control = rpart::rpart.control(minsplit = 20, cp = 0.01)) %T>% 
  rattle::fancyRpartPlot()

# Update missing survival with predicted survival
test_survived <- titanic %>% 
  filter(is.na(survived)) %>% 
  mutate(survived = predict(fit_survived, ., type = "class"))

# Recombine training and test datasets
titanic_dt <- bind_rows(train_survived, test_survived) %T>% 
  write_csv(here::here("titanic_dt.csv")) 
```


### Random Forest

[Trevor Stephen's series of titanic tutorials Part 5](http://trevorstephens.com/kaggle-titanic-tutorial/r-part-5-random-forests/)  calculated survival using random forest method

Random Forest method does not allow for missing values and requires factors for Survived and discrete variables

```{r eval=FALSE}
# Run random forest on training dataset with known survival
set.seed(415)
fit_survived <- randomForest::randomForest(survived ~ title_bins + age + pclass + sib_sp,
                data = train_survived,
                importance = TRUE,
                ntree = 2000) %T>% 
  randomForest::varImpPlot()

# Update missing survival with predicted survival
test_survived <- titanic %>% 
  filter(is.na(survived)) %>% 
  mutate(survived = predict(fit_survived, .))

# Recombine training and test datasets
titanic_rf <- bind_rows(train_survived, test_survived) %T>% 
  write_csv(here::here("titanic_rf.csv")) 

```

### Conditional inference trees

[Trevor Stephen's series of titanic tutorials Part 5](http://trevorstephens.com/kaggle-titanic-tutorial/r-part-5-random-forests/) calculated survival using conditional inference tree method

```{r eval=FALSE}
set.seed(415)

# Run Conditional inference trees on training dataset with known survival
fit_survived <- party::cforest(survived ~ title_bins + age + pclass + sib_sp,
                      data = train_survived,
                      controls = party::cforest_unbiased(ntree = 2000, mtry = 3))
                      
# Update missing survival with predicted survival
test_survived <- titanic %>% 
  filter(is.na(survived)) %>% 
  mutate(survived = predict(fit_survived, ., OOB = TRUE, type = "response"))

# Recombine training and test datasets
titanic_ci <- bind_rows(train_survived, test_survived) %T>% 
  write_csv(here::here("titanic_ci.csv")) 
```

## Kaggle submission

```{r eval=TRUE}
# Kaggle submission
submit <- test_survived %>%
  mutate(PassengerId = passenger_id,
                Survived = survived) %>% 
  select(PassengerId, Survived) %T>% 
  write_csv(here::here("kaggle_submission.csv")) 

submit %>% 
  tabyl(Survived) %>% 
  adorn_pct_formatting(digits = 1) 
```


| model | survived | died | kaggle_score |
|:----- |:-------- |:---- |:------------ |
| decision tree | 152 | 266 | 0.77511 |
| random forest | 151 | 267 | 0.76555 |
| conditional inference | 150 | 268 | 0.78468 |
  

